---
title: "Blood genesets"
author: "Junghyun Jung"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
```
### Essential genes

- A gene can be defined as **essential** when loss of its function compromises viability of the individual (for example, embryonic lethality) or results in profound loss of fitness.

- All the studies, whether in yeast, worms, mice or, more specifically, in humans, describe essential genes as located in critical places in the **core developmental, metabolic and signalling pathways**. Specifically, compared with other genes, essential genes are described as being broadly and strongly expressed and depleted in **expression quantitative trait loci (eQTLs).**^[[Human gene essentiality, Nat Rev Genet. 2018](https://www.nature.com/articles/nrg.2017.75)]

### Essential genes in Gene ontology terms
 - The [Gene ontology (GO)](https://en.wikipedia.org/wiki/Gene_ontology) provides a set of hierarchical controlled vocabulary split into 3 categories: **"Biological process"**, "Molecular function", "Cellular component".
 - Biological process terms, which involve over 150 genes, were used in this analysis. 
 - The GO terms were obtained from [Molecular Signatures Database.](http://software.broadinstitute.org/gsea/msigdb/index.jsp)

```{r echo=FALSE, fig.height=5, fig.width=3.5}
library(ggplot2)
load("data_c5bp.rda")
ggplot(data_c5bp,aes(data_c5bp$dev,data_c5bp$ratio,fill = data_c5bp$dev)) +
  geom_boxplot(notch = T)+
  theme_bw()+
  xlab("") +
  ggtitle("4,436 GO terms")+
  ylab("% essential gene")+
  theme(legend.position="none")+
  scale_fill_brewer(palette="Dark2")+
  theme(title = element_text(size = 15, color = "black"))+
  theme(axis.title.y = element_text(size = 12, color = "black"))+
  theme(axis.text.y = element_text(size = 9, colour = "black"))+
  theme(axis.text.x = element_text(size = 12, colour = "black"))

```

 - The GO terms associated with **development or morphogenesis contain essential genes at higher percentage**.

### Blood-related Gene Ontology terms

name | # genes|% essential|description |
-------------|-------|-----------|-------------------------------------------
[**Blood vessel morphogenesis**][2] | 364 |58.8 |The process in which the anatomical structures of blood vessels are generated and organized. The blood vessel is the vasculature carrying blood.
[Regulation of blood circulation][4]| 295 | 34.9 |Any process that modulates the frequency, rate or extent of blood circulation.	
[Regulation of blood pressure][5] | 169 |34.9 |Any process that modulates the force with which blood travels through the circulatory system. The process is controlled by a balance of processes that increase pressure and decrease pressure.
[**Circulatory system development**][3] | 788 |63.2| The process whose specific outcome is the progression of the circulatory system over time, from its formation to the mature structure. The circulatory system is the organ system that passes nutrients (such as amino acids and electrolytes), gases, hormones, **blood cells**, etc. to and from cells in the body to help fight diseases and help stabilize body temperature and pH to maintain homeostasis.


### conclusion
 - Circulatory system development and blood vessel morphogenesis are a core developmental process.
 
 - Considering the number of genes, **Blood vessel morphogenesis** is the most suitable for further analysis. 

[2]: http://amigo.geneontology.org/amigo/term/GO:0048514
[3]: http://amigo.geneontology.org/amigo/term/GO:0072359
[4]: http://amigo.geneontology.org/amigo/term/GO:1903522
[5]: http://amigo.geneontology.org/amigo/term/GO:0008217

